---
title: "covid 19"
author: "Christina Jilavdaryan"
date: "10/15/2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 
### This is contained in a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Question 1 - Is dying independent of sex


$H_0$: Dying is independent of sex.

$H_A$: Dying is dependent on sex.



Table of patients:
```{r}

# cross tabulation, patient state by gender

with(kRdata, table(Sex, result))

```

## We create two bar charts - one based on frequency and the other on percent, use ggplot commands to do this


```{r}
dd2 <- kRdata %>% group_by(result,Sex) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=result,y=count,fill=Sex))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=result,y=prcnt*100,fill=Sex)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```


Based on this ggplot, it appears that of those who died of Covid in South Korea, most were men. Women in South Korea appear to to have died of Covid at almost half the rate as men. The right bar of the graph(red) shows that woman died around 39%, while men(blue) died around 64% of times among those infected with the Covid virus.




  
## Numerical results



```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```

Based on this numerical table result it appears that of the people infected with Covid in South Korea, men died of Covid at almost twice the rate of women, with 64% of males infected with covid dying, while 36% of infected women died of Covid.

Overall, the table shows that very few people in South Korea were infected with Covid, and out of the people who were infected, with women in South Korea dying 1.4% of the times, and South Korean men dying 3% of the times showing that men died at twice the rate as women.


## Inferential results 


```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The p-value of 0.001 is low therefore we reject the null hypothesis. The death rate of covid in South Korea is dependent on sex.

## Conclusions

